---
title: "Conditional Aggregates"
output: html_notebook
---

```{r setup}
library(dplyr)
```

For this example, we'll use the mtcars data set.

```{r load_data}
data(mtcars)

head(mtcars)
```

It's straightforward to get aggregates using `group_by` and `summarize`. For example, we can get the average miles per gallon and weight by the numbers of cylinders in the car.

```{r simple_agg}
mtcars %>%
  group_by(cyl) %>%
  summarize(
    avg_mpg = mean(mpg),
    avg_wt = mean(wt)
  )
```

You can also imagine drilling down by grouping by two variables.

```{r groupby_2_var}
mtcars %>%
  group_by(cyl, gear) %>%
  summarize(
    avg_mpg = mean(mpg),
    avg_wt = mean(wt)
  )
```

But what if we want to get aggregates for all items in a group, but also add aggregate columns based on some condition? To get the conditional aggregate, just subset the variable inside the aggregate function.

```{r conditional_agg}
mtcars %>%
  group_by(cyl) %>%
  summarize(
    avg_mpg = mean(mpg),
    avg_wt = mean(wt),
    # Subset the variable inside the aggregate function
    avg_wt_3_gear = mean(wt[gear == 3]),
    avg_wt_4_5_gear = mean(wt[gear >= 4])
  )
```

