---
title: "Get a value of a different variable for each row of a dataframe based on a vector of variable names"
output: html_notebook
---

tl;dr: Use a matrix to index the dataframe.

Source: [r - How to extract different columns from each row of a data.frame?](https://stackoverflow.com/questions/25584039/how-to-extract-different-columns-from-each-row-of-a-data-frame)

## Motivation

You have a wide dataframe with consistently named columns and another column that can be used to determine which column to select.

## Create toy data

```{r make_data, echo = FALSE}
df <- data.frame(
  quarter = c(1, 2, 3, 4),
  q1_val = c(100, 200, 300, 400),
  q2_val = c(400, 100, 200, 300),
  q3_val = c(300, 400, 100, 200),
  q4_val = c(200, 300, 400, 100)
)

df
```

## Get variable names based on a variable

```{r get_variable_names}
var_names <- sprintf("q%d_val", df$quarter)

var_names
```

## Get numeric indices of the variable names

```{r get_var_name_indices}
var_name_indices <- match(var_names, colnames(df))

var_name_indices
```

## Use matrix indexing to get the values

```{r get_values}
df[cbind(seq_along(var_name_indices), var_name_indices)]
```